---
title: "Untitled"
author: '620128064'
date: "5/13/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(Hmisc)
library(dplyr)
library(ggplot2)
library(corrplot)
library(stringr)
library(lubridate)
library(tidyverse) 
library(caret)
library(ranger)
library(forcats)
library(stringr)
library(pROC)
library(gbm)
library(scales)

```


##Read information from dataset into dataframe
```{r}
crimeStat <- read_csv("/Users/sam/Downloads/atlcrime.csv")

##Create tables for data visualization, save table as text file for use when doing shiny application
table(crimeStat$crime)
x <- table(crimeStat$neighborhood)
write.table(x, file = "x.txt", sep = ",", quote = TRUE, row.names = F)
```

##Group the dataset by date, neighbourhood and crime and then calculate a count feature for occurence of crimes, add this column to dataset.
```{r}
crimeStat <- crimeStat %>% 
  group_by(date,  neighborhood, crime) %>% 
  dplyr::summarize(dailycount = n()) %>% 
  ungroup()
```

##Get the month, date, year and day of the week from each date, then add them into new columns with the respective names. Mutate these columns to a new dataframe, along with all previous columns. 
```{r}
crimeDaily <- crimeStat %>% 
  mutate(date = mdy(date), Month = month(date), Year = year(date),
         day = day(date),  
         weekday = weekdays(date, abbreviate=T), week = week(date)) %>% 
  mutate(weekday = factor(weekday, levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")))
names(crimeDaily)[1]='Date'
```

##Conduct a graphical visualization to see the amount of crimes per month.
```{r}
crimeDaily %>% 
  group_by(Month, crime) %>% 
  dplyr::summarize(count = sum(dailycount)) %>% 
  ggplot(aes(x = Month, y = count, col = crime)) +
  geom_point() +
  geom_line() +
  facet_wrap(.~crime, scales = 'free_y') + scale_x_continuous(breaks= pretty_breaks()) 
```

##Conduct graphical visualization to see the crime count per month
```{r}
  c <- crimeDaily %>% 
  group_by(Month, crime) %>% 
  dplyr::summarize(dailycount = sum(dailycount)) 
  ggplot(data = c, aes(x = Month, y = dailycount, col = crime)) + geom_point() + geom_line() +
    scale_x_continuous(breaks= pretty_breaks()) + ggtitle("Crimes per Month")
```

##Conduct graphical analysis to see the total crime counts per neighbourhood. Since there are 200+ neighbourhoods in dataset, provide a filter to see the most dangerous neighbourhoods.
```{r}
  crimeDaily %>% 
  group_by(neighborhood) %>%
  filter(neighborhood != "NA") %>%
  dplyr::summarize(total_count = sum(dailycount)) %>% 
  arrange(desc(total_count)) %>% 
  filter(total_count > 3967) %>% 
  ggplot(aes(x = fct_reorder(neighborhood, total_count), y = total_count)) +
  geom_bar(stat = 'identity') +coord_flip() +xlab('Neighborhood') +ylab('Total Crime Count') + ggtitle("Total Crime Count per Neighbourhood")
```

##Divide neighbourhoods into clusters since there are 200+ neighbourhoods in dataset. This would take far too long if fed to the model.
```{r}
cN <- crimeStat %>% 
  group_by(neighborhood,crime) %>% 
  dplyr::summarize(count = n()) %>% 
  drop_na() %>% 
  spread(crime,count,fill = 0) 
scaled = scale(cN[,2:7]) 
wss = 0 
for (i in 1:20) {
  km.out = kmeans(scaled, centers = i, nstart = 20, iter.max = 50)
  wss[i] = km.out$tot.withinss
}
plot(wss)
```
##Use the previous code to divide neighbourhoods into x amount of clusters, in this case we use 10, as that should be sufficient for what we are trying to dob
```{r}
km.out = kmeans(scaled, centers = 10, nstart = 20, iter.max = 50)
cN$cluster = km.out$cluster
nCluster <- cN %>% 
  select(neighborhood,cluster) 
nCluster$cluster = factor(nCluster$cluster)
```

##Preparing the data to so that count is initialized to 0 if a crime doesnt happen in a particular neighbourhood for a particular date.
```{r message=FALSE, warning=FALSE}
neighborhood = unique(crimeDaily$neighborhood[!is.na(crimeDaily$neighborhood)])
crime = unique(crimeDaily$crime)
Date = seq(ymd('2012-01-01'),ymd('2019-01-12'), by = '1 day')
prepare <- expand.grid(Date = Date,neighborhood = neighborhood,crime = crime) %>% 
  mutate(Month = month(Date), Year = year(Date),
         Day = day(Date),  
         Weekday = weekdays(Date, abbreviate=T), Week = week(Date)) %>% 
  mutate(Weekday = factor(Weekday, levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")))
crimeDaily <- prepare %>% 
  left_join(crimeDaily)
crimeDaily$dailycount[is.na(crimeDaily$dailycount)] = 0
table(crimeDaily$count)
proportion = crimeDaily %>% 
  group_by(crime) %>% 
  dplyr::summarize(prop = sum(dailycount != 0)/n())
proportion
```

##Create a variable indicating a boolean value if crime occurs or not in the different neighbourhoods.
```{r}
crimeDaily$crimeoccur = if_else(crimeDaily$dailycount == 0, 'No', 'Yes')
```

##Convert month and day to factors
```{r message=FALSE, warning=FALSE}
crimeDaily0 <- crimeDaily %>% 
  left_join(nCluster) %>% 
  mutate(Month = factor(Month), Day = factor(Day))
```
##Convert crimeoccur to a factor
```{r}
crimeDaily0$crimeoccur = factor(crimeDaily0$crimeoccur, levels = c('Yes','No'))
```

##Create the train and test datasets to be used by the model. Since the dataset is so vast, we will use a year worth of data to train the model, and dates from 2014 and larger for the testing dataset.
```{r}
trainHomicide <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime == 'HOMICIDE')
testHomicide <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime == 'HOMICIDE')

trainAggA <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime == 'AGG ASSAULT')
testAggA <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime == 'AGG ASSAULT')

trainAutoT <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime == 'AUTO THEFT')
testAutoT <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime == 'AUTO THEFT')

trainBur <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime == 'BURGLARY-NONRES'| crime == 'BURGLARY-RESIDENCE')
testBur <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime == 'BURGLARY-NONRES'| crime == 'BURGLARY-RESIDENCE')

trainLar <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime ==  'LARCENY-FROM VEHICLE' | crime == 'LARCENY-NON VEHICLE')
testLar <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime ==  'LARCENY-FROM VEHICLE' | crime == 'LARCENY-NON VEHICLE')

trainRape <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime ==  'RAPE')
testRape <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime == 'RAPE')


trainRobbery <- crimeDaily0 %>% filter(Date <ymd(20141001),Date >= ymd(20131001),crime ==  'ROBBERY-COMMERCIAL' | crime == 'ROBBERY-PEDESTRIAN' | crime == 'ROBBERY-RESIDENCE')
testRobbery <- crimeDaily0 %>% filter(Date >= ymd(20141001),crime ==  'ROBBERY-COMMERCIAL' | crime == 'ROBBERY-PEDESTRIAN' | crime == 'ROBBERY-RESIDENCE')
```



```{r}
glm_model <- glm(as.factor(crimeoccur) ~ neighborhood + Day + Month + Weekday,
                 data = trainHomicide, family = binomial)
```


```{r}
set.seed(32)
myFolds <- createFolds(trainHomicide$crimeoccur, k = 3)
myControl = trainControl(summaryFunction = twoClassSummary, classProbs = T,verboseIter = T,
                          savePredictions = F,
                          index = myFolds, trim = T, returnData = F)

formula_train = as.formula('as.factor(crimeoccur)~ Month + Day + Weekday')

```

##Create model using glmnet algorithm
```{r}
model_glmnet <- train(formula_train, trControl = myControl, tuneGrid = expand.grid(alpha = c(0,0.1,1), lambda = c(0,0.5,1)),
                      method = 'glmnet',metric = 'ROC', data = trainHomicide)
```

##Visualize results
```{r}
plot(model_glmnet)
print(model_glmnet)
```


##Train model for each crime type
```{r}
glmnet_model_homicide = train(crimeoccur ~ cluster + Day + Month + Weekday,
                     data = trainHomicide,
                     metric = 'ROC',
                     method = "glmnet",
                     tuneGrid = expand.grid(
                       alpha = 1,
                       lambda = 0
                     ),
                     trControl = myControl)
```

```{r}
glmnet_model_aggA = train(crimeoccur ~ cluster + Day + Month + Weekday,
                              data = trainAggA,
                              metric = 'ROC',
                              method = "glmnet",
                              tuneGrid = expand.grid(
                                alpha = 1,
                                lambda = 0
                              ),
                              trControl = myControl)
```

```{r}
glmnet_model_AutoT = train(crimeoccur ~ cluster + Day + Month + Weekday,
                             data = trainAutoT,
                             metric = 'ROC',
                             method = "glmnet",
                             tuneGrid = expand.grid(
                               alpha = 1,
                               lambda = 0
                             ),
                             trControl = myControl)
```                            

```{r}
glmnet_model_Bur = train(crimeoccur ~ cluster + Day + Month + Weekday,
                             data = trainBur,
                             metric = 'ROC',
                             method = "glmnet",
                             tuneGrid = expand.grid(
                               alpha = 1,
                               lambda = 0
                             ),
                             trControl = myControl)
```

```{r}
glmnet_model_Lar = train(crimeoccur ~ cluster + Day + Month + Weekday,
                             data = trainLar,
                             metric = 'ROC',
                             method = "glmnet",
                             tuneGrid = expand.grid(
                               alpha = 1,
                               lambda = 0
                             ),
                             trControl = myControl)
```

```{r}
glmnet_model_rape = train(crimeoccur ~ cluster + Day + Month + Weekday,
                             data = trainRape,
                             metric = 'ROC',
                             method = "glmnet",
                             tuneGrid = expand.grid(
                               alpha = 1,
                               lambda = 0
                             ),
                             trControl = myControl)
```

```{r}
glmnet_model_Robbery = train(crimeoccur ~ cluster + Day + Month + Weekday,
                             data = trainRobbery,
                             metric = 'ROC',
                             method = "glmnet",
                             tuneGrid = expand.grid(
                               alpha = 1,
                               lambda = 0
                             ),
                             trControl = myControl)
```


##Use generalization to visualize results
```{r}
roc_obj <- roc(testHomicide$crimeoccur, predict(glmnet_model_homicide, testHomicide, type = 'prob')$Yes)
auc(roc_obj)
```

```{r}
roc_obj <- roc(testAggA$crimeoccur, predict(glmnet_model_aggA, testAggA, type = 'prob')$Yes)
auc(roc_obj)
```
```{r}
roc_obj <- roc(testAutoT$crimeoccur, predict(glmnet_model_AutoT, testAutoT, type = 'prob')$Yes)
auc(roc_obj)
```
```{r}
roc_obj <- roc(testBur$crimeoccur, predict(glmnet_model_Bur, testBur, type = 'prob')$Yes)
auc(roc_obj)
plot(roc_obj)
```

```{r}
roc_obj <- roc(testRape$crimeoccur, predict(glmnet_model_rape, testRape, type = 'prob')$Yes)
auc(roc_obj)
plot(roc_obj)
```
```{r}
roc_obj <- roc(testLar$crimeoccur, predict(glmnet_model_Lar, testLar, type = 'prob')$Yes)
auc(roc_obj)
```

```{r}
roc_obj <- roc(testRobbery$crimeoccur, predict(glmnet_model_Robbery, testRobbery, type = 'prob')$Yes)
auc(roc_obj)
```

##Make a random prediction to ensure model is working
```{r}
prediction2 <-  predict(glmnet_model_aggA, data.frame(cluster = "6", Day = "17", Month = "2", Weekday = "Mon"), type='prob')
prediction2$Yes * 100
```

##Split dataset into locations, to get the lattitude and longitude values of each neighbourhood
```{r}
loc <- read_csv("/Users/sam/Downloads/atlcrime.csv")
locations <- loc %>%
  group_by(neighborhood,lat,long) %>% 
  dplyr::summarize(count = n()) %>% 
  drop_na() 
```

```{r}
lattitude <- locations %>%
  group_by(neighborhood) %>%
  dplyr::summarize(lat= mean(as.numeric(lat))) 
```

```{r}
longitude <- locations %>%
  group_by(neighborhood) %>%
  dplyr::summarize(longg = mean(as.numeric(long)))
```

```{r}
totalLoc <- lattitude %>%
  left_join(longitude)
```

##More visualizations for dataset, to see crimes per month for particular years.
```{r}
c2016 <- crimeDaily %>% 
  group_by(Month, Year, crime) %>% 
  filter(Year == 2016) %>%
  dplyr::summarize(dailycount = sum(dailycount)) 
ggplot(data = c2016, aes(x = Month, y = dailycount, col = crime)) + geom_point() + geom_line() +
scale_x_continuous(breaks= pretty_breaks()) + ggtitle("Crimes per Month for 2016")
```

```{r}
c2015 <- crimeDaily %>% 
  group_by(Month, Year, crime) %>% 
  filter(Year == 2015) %>%
  dplyr::summarize(dailycount = sum(dailycount)) 
ggplot(data = c2015, aes(x = Month, y = dailycount, col = crime)) + geom_point() + geom_line() +
scale_x_continuous(breaks= pretty_breaks()) + ggtitle("Crimes per Month for 2015")
```

```{r}
c2014 <- crimeDaily %>% 
  group_by(Month, Year, crime) %>% 
  filter(Year == 2014) %>%
  dplyr::summarize(dailycount = sum(dailycount)) 
ggplot(data = c2014, aes(x = Month, y = dailycount, col = crime)) + geom_point() + geom_line() +
scale_x_continuous(breaks= pretty_breaks()) + ggtitle("Crimes per Month for 2014")
```


##More visualizations to see the neighbourhoods with the most crime levels for particular years.
```{r}
  top2014crimeDaily <- crimeDaily %>% 
  group_by(neighborhood) %>%
  dplyr::filter(neighborhood != "NA") %>%
  dplyr::summarize(total_count = sum(dailycount)) %>%
  arrange(desc(total_count)) %>% 
  dplyr::filter(total_count > 4000) 
```

```{r}
 top2015crimeDaily <- crimeDaily %>% 
   group_by(neighborhood)%>%
    dplyr::filter(Year==2015) %>%
    dplyr::filter(neighborhood!="NA") %>%
    dplyr::summarize(total_count = sum(dailycount)) %>%
    arrange(desc(total_count)) %>% 
    dplyr::filter(total_count > 700) 
```

```{r}
  top2016crimeDaily <- crimeDaily %>% 
  group_by(neighborhood) %>%
  dplyr::filter(Year==2016) %>%
  dplyr::filter(neighborhood != "NA") %>%
  dplyr::summarize(total_count = sum(dailycount)) %>%
  arrange(desc(total_count)) %>% 
  dplyr::filter(total_count > 680) 
```

```{r}
ggplot(data = top2014crimeDaily, aes(x = neighborhood, y=total_count))+ geom_bar(stat = 'identity') +coord_flip() +xlab('Neighborhood') +ylab('Total Crime Count') + ggtitle("Total Crime Count per Neighbourhood")
```


